XXX

### 2012-03-25 paper/engNERwiki.tex
Hungarian resz megirva

While the correspondence described above holds for most classes in the XXX sub-hierarchies, there are some exceptions. For instance, the class \texttt{SportsLeague} is part of the \texttt{Organisation} sub-hierarchy, but according to the CoNLL annotation scheme, they are tagged as MISC. Similarly, the class \texttt{SportsTeamSeason} is under \texttt{Event}, by which account it would be classified as MISC; however, ConLL does not recognize such entities. To account for these misclassification, as well as the miscellaneous classes outside the main hierarchies, we have introduced a mapping mechanism. We created a file of DBpedia class--NE category mappings. Whenever an entity is evaluated, we look up its class, as well as the ancestors of its class, in the mapping, and assign to the entity the category of the class that matches it most closely -- i.e.~the category of the entity's class, if it is in the list, or that of its closest ancestor. XXX!!! If no match is found, the entity is tagged with \texttt{O}. Since we take advantage of the inheritance hierarchy, the mapping list remains short: it contains only the root classes of the main sub-hierarchies, exceptions of the like XXX mentioned above and the various classes that belong to the MISC category according to the CoNLL annotation guideline.

As of version 3.7, the DBpedia ontology allows multiple superclasses, making the hierarchy a directed acyclic graph (CITE). This introduces the problem of selecting the right superclass, and hence, CoNLL mapping for classes with more than one parent. In version 3.7, the only such class is \textit{Library}, which can be traced back to both \textit{Place} and \textit{Organisation}. There are two ways to tackle this problem. The first is to follow the DBpedia guidelines and only retain the first parent. The second method is manually deciding the right NE mapping for the ambiguous XXX class and adding it to the mapping file. The first approach, unfortunately, already fails to assign the correct CoNLL NE tag to our only example. The reason behind this failure is that multiple parents in an ontology show an inherent ambiguity in the class in question: it is entirely natural to think of a library as a location as well as a public organization. DBpedia's guidelines prefers Organisation; while ConLL's one refers it as a Location. Therefore, only the second, manual approach ensures correct tagging. XXX

@@

\subsection{Extracting and Filtering Sentences}

A NER corpus consists of sentences that accurately tag NEs occurring in them. In order to automatically prepare such sentences, two tasks need to be performed: identifying the entities in the sentence and tagging them with the correct tag. Sentences, for which accurate tagging could not be accomplished, must be removed from the corpus. In this section, we describe how these steps were implemented, as well as the post-filtering we employed to dispose of correctly tagged, but unusable training sentences.

\subsubsection{Identifying the entities}
% TODO: identification and tagging happens in parallel, else the recall oriented approach cannot be explained.

@@

However, not all entities are marked with links. Authors may neglect to add XXX references to their text. Furthermore, in a Wikipedia article XXX, typically only the first occurrence of a particular entity is linked to the corresponding page. Subsequent mentions are unmarked and often incomplete -- e.g.~family names are used instead of full names. We devised XXX two methods to account for unmarked entities: one that emphasizes XXX precision, and the other, recall. XXX

% Or NPP

The precision-oriented approach seeks to minimize the risk of words getting incorrectly tagged XXX in the corpus. Names and their parts mostly start with capital letter, thus 
%A short study of the CoNLL annotation guidelines reveals that the words in the entities of interest to us all start with capital letter. According to this method XXX, 
all sentences that contain capitalized words, but are not part of the anchor text of a link, XXX were discarded. The rationale behind this approach is that the sheer size of Wikipedia guarantees that we end up with a sizeable corpus even with such a strict policy. XXX

The recall-oriented method tries to associate unmarked entities with entities already identified on the same page. Only if the attempt fails is the sentence that contains the words discarded. 
%Ez utóbbi biztosan jó mondat?
This approach aims to generate more sentences with as little loss of precision as possible.

The algorithm maintains a list of entities already indentified on the page. The list includes:
\begin{itemize}
\item The page title; % if it is in DBpedia
\item Title of redirect pages that target the current page; % TODO
\item Entities marked with links in the text thus far.
\end{itemize}

@@

%\subsubsection{Tagging}

%TODO: merge with Entity Identification
%megtörtént

@@

Strictly speaking, our original assumption of equating Wikipedia articles with NEs is not valid: many pages describe common nouns (Book, Aircraft), numbers (1, 42), currencies (Euro, Dollar) or other concepts that fall outside the scope of CoNLL tagset or of NER in general. In this section, we describe the measures taken to avoid errors introduced by these false entities. We shall modify the two approaches -- precision and recall-oriented -- accordingly.

@@

Links to number and common noun pages are handled differently by the two approaches. The precision-oriented method treats them similarly to links to unknown entities and discards the sentences in which they occur. The recall-oriented approach tries to distinguish them from links to entities so that otherwise valid sentences can be retained XXX. The method is as follows XXX.
\begin{itemize}
\item Common nouns are filtered by POS category: links that contain no proper nouns are ignored.
\item Numbers behave more like proper nouns in that they can refer to entities, such as 737 to the Boeing 737. Therefore, links to number pages are filtered by the surface form of the page title.
\end{itemize}

@@

% 4. Test sentence extraction?
%   a. unknown links, no unlinked names
%   b. might be good for quality evaluation / recursive tagging (next paper?)
% 5. teaching the original DBpedia categories?

In Wikipedia the nationalities (e.g.~American and Australian) links to the corresponding places (America and Australia, respectively). However, in CoNLL scheme these types of entities take parts of MISC category. XXX Postprocessing: list look-up (nationalities, adjective forms of person who live in the corresponding country); if the name is POS-tagged as adjective, then it is not a NE

@@

\subsection{Tools}

Although Hungarian is reckoned to be a less resourced language, and it is not supperted in NLTK, several high quality language processing tools have been developed for Hungarian in recent years. For tokenization and sentence segmentation we used an in-house statistical tool tailored for Hungarian. It has been trained on the largest manually annotated Hungarian corpus \cite{Csendes:04}, and it handles the peculiarities of Hungarian ortography, such as the periods placed after numbers in date expressions. Lemmatization was performed by \texttt{hunmorph} \cite{Tron:05} and \texttt{hundisambig}, latter being an in-house disambiguator to select the right analysis based on the word context.

@@

The Hungarian morphological analyzer outputs KR-codes \cite{Kornai:04}. In addition to the POS category, the KR-code also includes inflection information, such as the number, person and case XXX, which makes it much better suited to agglutinative languages than Penn Treebank POS tags. One shortcoming of the KR code is that it does not differenciate between common and proper nouns. Since in Hungarian only proper nouns are capitalized, we can usually decide whether a noun is proper or not based on the initial letter. However, this rule cannot be used if the noun is at the beginning of a sentence, as words at the beginning of sentences are always capitalized. Hence, all sentences that begin with nouns have been removed from the corpus.

\subsection{Data Availability}

For some of the bigger languages, DBpedia has internationalized "chapters" -- data dumps based on the Wikipedia of that particular language. Hungarian, being a relatively small language XXX, does not have a separate chapter. Instead, the Hungarian entity list comprises of the pages in the English list that have their equivalents in the Hungarian Wikipedia. Two consequences follow.

First, in order to identify which pages denote entities in the Hungarian Wikipedia, an additional step is required, in which the Hungarian equivalents of the English pages are added to the entity list. The English titles are retained, as due to the relatively small size of the Hungarian Wikipedia, in-article links sometimes point to English articles.

@@

\subsection{NER Conventions}

% Eszter
The largest (?, the only?, most cited?) NER corpus in Hungarian is the Szeged NER corpus, which will be described by Eszter. :) The Szeged corpus uses the same named entity categories as ConLL; however, the convention of what constitutes an entity is slightly different.

The Szeged NER guideline relies heavily on the rules of capitalization to decide which words should be marked as named entities. As mentioned earlier, in Hungarian, apart from the first word in the sentence, only proper nouns are capitalized. The following concepts are not proper nouns in Hungarian and thus are not considered named entities according to the Szeged guideline:
\begin{itemize}
\item Names of languages and ethnic groups
\item Words created from proper nouns through derivation; most prominently
\item Adjectival form of location names
\item Months
\end{itemize}

The strong correlation between capitalization and named entities makes recognition of entities in text relatively simple: a word is part of an entity if it is capitalized.
% Eszter
a tulajdonnevekhez kötőjellel kapcsolódó köznevek kérdésesek (pl. SAP-szoftvercég, Matáv-csoport)

% Ezt hová?
most van: 102 millió szó a disambiguation pages-zel együtt
az angol: 1,7 milliárd

@@

@@

>>>>

While the correspondence described above holds for most classes in the XXX sub-hierarchies, there are some exceptions. For instance, the class \texttt{SportsLeague} is part of the \texttt{Organisation} sub-hierarchy, but according to the CoNLL annotation scheme, they are tagged as MISC. Similarly, the class \texttt{SportsTeamSeason} is under \texttt{Event}, by which account it would be classified as MISC; however, CoNLL does not recognize such entities.

To account for these misclassification, as well as the miscellaneous classes outside the main hierarchies, we have introduced a mapping mechanism. We created a file of DBpedia class--NE category mappings. Whenever an entity is evaluated, we look up its class, as well as the ancestors of its class, in the mapping, and assign the category of the class that matches it most closely to the entity. If no match is found, the entity is tagged with \texttt{O}. Since we take advantage of the inheritance hierarchy, the mapping list remains short: it contains only the root classes of the main hierarchies, exceptions like those mentioned above, and the various classes that belong to the MISC category according to the CoNLL annotation guideline.

As of version 3.7, the DBpedia ontology allows multiple superclasses, making the hierarchy a directed acyclic graph (CITE). This introduces the problem of selecting the right superclass, and hence, CoNLL mapping for classes with more than one parent. In version 3.7, the only such class is \textit{Library}, which can be traced back to both \textit{Place} and \textit{Organisation}. There are two ways to tackle this problem. The first is to follow the DBpedia guidelines and only retain the first parent. The second method is manually deciding the right NE mapping for the ambiguous XXX class and adding it to the mapping file.

The first approach, unfortunately, already fails to assign the correct ConLL NE tag to our only example. The reason behind this failure is that multiple parents in an ontology show an inherent ambiguity in the class in question: it is entirely natural to think of a library as a location as well as a public organization. DBpedia's guidelines prefers Organisation, while ConLL's one refers it as a Location. Therefore, only the second, manual approach ensures correct tagging. XXX

@@

Using the mapping described above, we compile an entity list that contains all entities in DBpedia tagged with the appropriate XXX CoNLL NE category.

\subsection{Named Entity Labeling}

A NER corpus consists of sentences where NEs are accurately tagged. In order to automatically prepare such sentences, two tasks need to be performed: identifying the entities in the sentence and tagging them with the correct tag. Sentences, for which accurate tagging could not be accomplished, must be removed from the corpus.

Our approach to entity identification relies on Wikipedia cross-references found in the article text. We assume that individual Wikipedia articles describe named entities. A link to an article can then be perceived as a mapping that identifies its anchor text with a particular entity.

Discovered entities are tagged with the CoNLL category assigned to them in the entity list extracted from DBpedia. If the link target is not in the entity list, or the link points to a disambiguation page, we cannot determine the type of the entity and tag it as \texttt{UNK}. Links to redirect pages are resolved to point instead to the redirect target, after which they are handled as regular cross-references.

There are several

\subsubsection{Non-entity Links}

@@

%Strictly speaking, our original assumption of equating Wikipedia articles with named entities is not valid: many pages describe common nouns (Book, Aircraft), numbers (1, 42), currencies (Euro, Dollar) or other concepts that fall outside the scope of NER.

\begin{description}
\item[Common nouns] Links to common nouns are filtered by POS tags; if a link contains no \texttt{NNP}s, it is ignored.
\item[Dates] Wikipedia has pages for years, months and concrete dates, therefore our algorithm might mistake them for entities. To prevent this problem, we compiled a list of calendar-related entities and added them to the main entity list tagged with the CoNLL category \texttt{O}.
\item[Lowercase links] 
\end{description}

\subsubsection{Unmarked Entities}

\subsubsection{Special Cases}

Not all entities are marked with links. Authors may neglect to add references to their text. Furthermore, in a Wikipedia article, typically only the first occurrence of a particular entity is linked to the corresponding page. Subsequent mentions are unmarked and often incomplete -- e.g.~family names are used instead of full names. To account for such entities,

The recall-oriented method tries to associate unmarked entities with entities already identified on the same page. Only if the attempt fails is the sentence that contains the words discarded. 
This approach aims to generate more sentences with as little loss of precision as possible.

The algorithm maintains a list of entities already indentified on the page. The list includes:
\begin{itemize}
\item The page title, if we know its category;
\item Title of redirect pages that target the current page; % TODO
\item Entities marked with links in the text thus far.
\end{itemize}

@@

@@

%Strictly speaking, our original assumption of equating Wikipedia articles with NEs is not valid: many pages describe common nouns (Book, Aircraft), numbers (1, 42), currencies (Euro, Dollar) or other concepts that fall outside the scope of CoNLL tagset or of NER in general. In this section, we describe the measures taken to avoid errors introduced by these false entities. We shall modify the two approaches -- precision and recall-oriented -- accordingly.

@@

Links to number and common noun pages are handled differently by the two approaches. The precision-oriented method treats them similarly to links to unknown entities and discards the sentences in which they occur. The recall-oriented approach tries to distinguish them from links to entities so that otherwise valid sentences can be retained XXX. The method is as follows XXX.

\begin{itemize}
\item Common nouns are filtered by POS category: links that contain no proper nouns are ignored.
\item Numbers behave more like proper nouns in that they can refer to entities, such as 737 to the Boeing 737. Therefore, links to number pages are filtered by the surface form of the page title.
\end{itemize}

@@

In Wikipedia the nationalities (e.g. American and Australian) links to the corresponding places (America and Australia, respectively). However, in CoNLL scheme these types of entities take parts of MISC category. XXX Postprocessing: list look-up (nationalities, adjective forms of person who live in the corresponding country); if the name is POS-tagged as adjective, then it is not a NE

@@

\subsection{Parsing the Hungarian Wikipedia}

Although Hungarian is reckoned to be a less resourced language, and it is not supperted in NLTK, several high quality language processing tools have been developed for Hungarian in recent years. For tokenization and sentence segmentation we used an in-house statistical tool tailored for Hungarian. It has been trained on the largest manually annotated Hungarian corpus \cite{Csendes:04}, and it handles the peculiarities of Hungarian ortography, such as the periods placed after numbers in date expressions. Lemmatization was performed by \texttt{hunmorph} \cite{Tron:05} and \texttt{hundisambig}, the latter being an in-house disambiguator to select the right analysis based on the word context.

@@

Hungarian is called a synthetic language: for the most part, it expresses grammatical elements in a single word form using affixes, as opposed to isolating languages (as English), which tend to employ separate words, e.g.~prepositions, pronouns, auxiliaries, for expressing grammatical phenomena. The Hungarian morphological analyzer outputs KR-codes \cite{Kornai:04}, which, in addition to the POS category, also includes inflectional information, which makes it much better suited to agglutinative languages, than Penn Treebank POS tags. One shortcoming of the KR-code is that it does not differenciate between common and proper nouns. Since in Hungarian only proper nouns are capitalized, we can usually decide whether a noun is proper or not based on the initial letter. However, this rule cannot be used if the noun is at the beginning of a sentence, as words at the beginning of sentences are always capitalized. Hence, all sentences that begin with nouns have been removed from the corpus.

\subsection{Named Entity Labeling in Hungarian}

For some of the well-resourced languages, DBpedia has internationalized "chapters" -- data dumps based on the Wikipedia of those particular languages. Hungarian does not have a separate chapter. Instead, the Hungarian entity list comprises of the pages in the English list that have their equivalents in the Hungarian Wikipedia. Two consequences follow.

First, in order to identify which pages denote entities in the Hungarian Wikipedia, an additional step is required, in which the Hungarian equivalents of the English pages are added to the entity list. The English titles are retained, as due to the medium size of the Hungarian Wikipedia, in-article links sometimes point to English articles.

@@

To annotate our corpus with NE-tags, we chose to follow the annotation guidelines of the largest human-annotated NER corpus for Hungarian, Szeged NER corpus \cite{Szarvas:06}. It is quite similar to the CoNLL standards: it contains business newswire texts, it comprises ca.~200,000 tokens, and the text has been annotated with NE class labels in line with the CoNLL annotation scheme. However, the convention of what constitutes a NE is slightly different for Hungarian.

\subsubsection{Special cases}

The Szeged NER guideline relies heavily on the rules of capitalization to decide which words should be marked as NEs. As mentioned earlier, in Hungarian, apart from the first word in the sentence, only proper nouns are capitalized. The following concepts are not proper nouns in Hungarian, and thus are not considered as NEs according to the Szeged NER guideline:

\begin{itemize}
\item names of languages, nationalities, religions, poltical ideologies;
\item adjectives derived from a NE;
\item names of months, days, holidays;
\item names of special events and wars.
\end{itemize}

There is an other special case in Hungarian: opposite to English, the number of compund words is quite large in Hungarian, and NEs can also be subjects of compounding. In this case the common noun following the NE is joint with a hyphen, so they constitute one token. However, the joint common noun can modify the original sense of NE, depending on the semantics of the common noun. For example: in the compound \textit{SAP-szoftvercég} ['SAP software company'] the common noun does not change the labeling, but in the case of the compound \textit{WorldCom-botrány} ['WorldCom scandal'] the NE tag changes from ORG to O.

Inflections of acronyms and foreign names ending with non-pronounced vowel have similar surface forms to the aforementioned compounds, e.g.~\textit{MTI-t} ['MTI'+ACC], \textit{Shakespeare-rel} ['with Shakespeare']. It is important to distinguish these types of hyphenated NEs, because inflections do not change NE labeling in contrast to some type of compounds. Due to the surface similarity of these phenomena the solution of this problem is not obvious, and needs more investigation.

% Ezt hová?
%most van: 102 millió szó a disambiguation pages-zel együtt
%az angol: 1,7 milliárd

@@

Although two human-annotated NER corpora exist for Hungarian, we chose the Szeged NER corpus \cite{}, which follows the CoNLL annotation scheme. It contains ca.

The strong correlation between capitalization and named entities makes recognition of entities in text relatively simple: a word is part of an entity if it is capitalized.

@@

\bibitem[\protect\citename{Szarvas \bgroup et al.\egroup}2006]{Szarvas:06}
György Szarvas, Richárd Farkas, András Kocsor.
\newblock 2006.
\newblock A highly accurate Named Entity corpus for Hungarian. 
\newblock In: {\em Proceedings of International Conference on Language Resources and Evaluation}.
